---
title: "NFLData"
author: "Navin Chandradat, Minhaz Khan, Vincent La, Truman Zheng, Stephen Pereira"
date: "11/13/2018"
output: html_document
---

```{r}
library(tidyverse)
```


First to read in and display our data:

```{r}
my_url = "http://www.utsc.utoronto.ca/%7ebutler/c32/qbrating.txt"

qbrating = read.table(my_url, head=T)
qbrating
```

Looking at the statistics in the data frame and using our prior knowledge that the qbrating was calculated over the whole season we will only consider the season totals when running our analysis.


Steps:
-Introduction
-Adjust the units of the data before any analysis can be done (b/c its per game not per season or vice versa) *If the stat is per game multiply it by 16 to get the value for the whole season because an NFL season consists of 16 games*
-Potential Residual Plot
-Run a multiple regression
-See which variables are significant
-Run the regression again without the insignificant variables
-Summary/Conclusion



test - Vincent
test - Stephen hello world

##Plot of Stats against Rating
```{r}
qbrating %>% gather(stats,x, c(completions:sacks)) %>% 
  ggplot(aes(x=x,y=rating,color = player))+geom_point()+facet_wrap(~stats,scales = "free")
```
