---
title: "NFLData"
author: "Navin Chandradat, Minhaz Khan, Vincent La, Truman Zheng, Stephen Pereira"
date: "11/13/2018"
output: html_document
---


# Going deep: What Are the Makings of a Prolific Passer in the NFL? 

## Introduction

In any professional sports league (here, being the NFL), one of the things us fans really like to do is rank players in terms of skill. 
One of the most common ways for fans to make an argument for one player being better than another is using advanced statistics, more 
specifically, comparing their player efficiency ratings (PER). This efficiency rating is calculated using the player's per game 
statistics but obviously some statistics should be weighted more than others (i.e. touchdowns should be worth more than number of 
passing attempts.). Our goal in this report is to figure out how exactly the NFL calculates player efficiency rating through the use of techniques learned within this course.


The packages we will need for our analysis:
```{r}
library(tidyverse)
library(broom)
```


First to read in and display our data:
```{r}
my_url = "http://www.utsc.utoronto.ca/%7ebutler/c32/qbrating.txt"

qbrating = read.table(my_url, head=T)
qbrating
qbrating2 = qbrating %>% separate(long, c("long", "went.for.TD"), 2)
qbrating2
qbrating2$long <- as.integer(as.character(qbrating2$long))
qbrating2
```
Here, our data is a 49 rows by 20 columns data set with categories rankings, player names, teams and various player statistics (i.e. completions, attempts).

Looking at the statistics in the data frame and using our prior knowledge that the qbrating was calculated over the whole season we will only consider the season totals when running our analysis. There for attempts.game and yards.per.game will be omitted because they are per game stats.


##Plot of Stats against Rating
```{r}
qbrating2 %>% gather(stats,x, c(completions:percent.complete, yards:average.yards, td:long, twenty.plus:sacks)) %>% 
  ggplot(aes(x=x,y=rating))+geom_point()+facet_wrap(~stats,scales = "free")
```
Here we plotted the stats against of rating of each player to get some visuals of the data and also get some potential pattern between each individual player statistic and rating. Here we see that for each category there seems to be some sort of linear trends, so maybe a linear model would be best when it comes to model fitting.


## Attempt to do residuals.
```{r}
.remove = qbrating2[,-(1:4)]
model2 = lm(rating~completions+attempts+percent.complete+attempts.game+yards+average.yards+yards.per.game+td+int+first.down+first.down.percent+long+twenty.plus+forty.plus+sacks, data=.remove)
g.1 = ggplot(model2, aes(y=.resid,x=.fitted))+geom_point()
g.1
# residuals = resid(model2)
# plot(residuals)
```


## Attempt to do regression:

```{r}
NFL_reg = lm(rating~completions+attempts+percent.complete+yards+average.yards+td+int+first.down+first.down.percent+long+twenty.plus+forty.plus+sacks, data=qbrating2)
summary(NFL_reg)

tidy(NFL_reg) %>% arrange(p.value)
```
Only average.yards, td, int, first.down.percent have p-values below 0.05. What if we remove forty.plus (highest p-value)?

```{r}
NFL_reg2 = lm(rating~completions+attempts+percent.complete+yards+average.yards+td+int+first.down+first.down.percent+long+twenty.plus+sacks, data=qbrating2)
summary(NFL_reg2)

tidy(NFL_reg2) %>% arrange(p.value)
```
The value for R-squared (0.9253) does not change by much, although we do see a shift in p-values. What if we remove twenty.plus now?

```{r}
NFL_reg3 = lm(rating~completions+attempts+percent.complete+yards+average.yards+td+int+first.down+first.down.percent+long+sacks, data=qbrating2)
summary(NFL_reg3)

tidy(NFL_reg3) %>% arrange(p.value)
```
Still not a large change in R-squared. We remove completions:

```{r}
NFL_reg4 = lm(rating~attempts+percent.complete+yards+average.yards+td+int+first.down+first.down.percent+long+sacks, data=qbrating2)
summary(NFL_reg4)

tidy(NFL_reg4) %>% arrange(p.value)
```
We also drop long:

```{r}
NFL_reg5 = lm(rating~attempts+percent.complete+yards+average.yards+td+int+first.down+first.down.percent+sacks, data=qbrating2)
summary(NFL_reg5)

tidy(NFL_reg5) %>% arrange(p.value)
```
Dropping sacks:

```{r}
NFL_reg6 = lm(rating~attempts+percent.complete+yards+average.yards+td+int+first.down+first.down.percent, data=qbrating2)
summary(NFL_reg6)

tidy(NFL_reg6) %>% arrange(p.value)
```
Despite removing many variables, we do not see a significant change in our R-squared. Let us see what happens when we remove first.down:

```{r}
NFL_reg7 = lm(rating~attempts+percent.complete+yards+average.yards+td+int+first.down.percent, data=qbrating2)
summary(NFL_reg7)

tidy(NFL_reg7) %>% arrange(p.value)
```
After removing first.down there does not seem to be much change in the R-squared value and we see that what we have left are now only the variables for which P-value are less than 0.05, which means that all those variable are now significant and this is where we stop the elimination process.


## Residuals for model 7
```{r}
g.2 = ggplot(NFL_reg7, aes(y=.resid,x=.fitted))+geom_point()
g.2
```

```{r}
g3=augment(NFL_reg7, qbrating2) %>% 
  gather(stats,x, c(completions:percent.complete, yards, td:long, twenty.plus:sacks)) %>%
  ggplot(aes(x=x,y=.resid))+geom_point()+facet_wrap(~stats,scales = "free")
g3
```
Here, we see that none of the plot showed any sort of pattern, could say there is a cluster for percent.complete and first.down.percent however that is just the nature of the data points, so is not a concern. So from this we can confirm the validity of our model.


### Side note
Also noting, the actual formula takes into account attempts, number of completions, yards, touchdowns and interceptions. Our analysis did come very close by including attempts, completion percentage (completions divided by attempts), yards, touchdowns and interceptions and yards per pass (yards divided by attempts), 
but our model also included first.down.percentage which isn't included in the actual formula.    

## Conclusion

After careful analysis we can see that not all variables provided in the dataset were important in predicting the quarterbacks' 
efficiency ratings. If we look at model 7, it is the best in terms of how good our R-squared value is and the fact that whenever we
drop another variable from the model, R-squared takes a big hit. The significant variables in model 7 are the following: attempts,
percent.complete, yards, average.yards, td, int, and first.down.percent. Intuitively, this makes sense. The model takes into account
both the offensive (in touchdowns and passing, which is expressed by yards and average.yards) and defensive (interceptions)
aspects of the game, while taking into account consistent, smart play in percent.complete (i.e. successfully catching the ball so the 
enemy team does not get it). The other statistics are heavily influenced by these 7 so they can be dropped since they are essentially 
side effects of the others. Hence, the most important variables are the ones expressed within model 7. 
